Preregistration analysis, Cost 2 study

### 2019-05-09 Cost2/Cost2_prereg.Rmd
slight changes to plots

#### 4.2.	Prospect acceptance rates will differ across four between-subject conditions, in which the delays associated with rewards (a) are unfilled, (b) include a cognitive effort requirement, (c) include a physical effort requirement, or (d) require a trivial level of physical effort but have matched visual stimuli to the physical effort condition.
  **4.2.1.**	In the cognitive effort condition, overall acceptance rates will be greater than in the unfilled-delay condition.

@@

#### 4.3.	Choices will be well fit by a computational model in which the subjective opportunity cost of time is free to vary across the four between-subject conditions.
  **4.3.1.**	Participants will display stable preferences, meaning that the reward amounts they accept in a given timing condition will be similar throughout the experiment.

@@

@@

``` {r Load Data, echo=FALSE}
### data loading and cleaning up

# load cognitive effort group
setwd('./data/')
temp = list.files(pattern="cogTask")
cogDataLong = lapply(temp, read.csv, header = FALSE)

# load wait group
temp = list.files(pattern="wait")
waitDataLong = lapply(temp, read.csv, header = FALSE)

# load physical group
temp = list.files(pattern="phys")
physDataLong = lapply(temp, read.csv, header = FALSE)

# load pheasysical group
temp = list.files(pattern="pheasy")
easyDataLong = lapply(temp, read.csv, header = FALSE)

### cleaning data
# this will generate a simple list with only info about trials that participants quit or completed
# this is because cognitive logs also include within-trial task-specific information
# for the wait condition, it will get rid of the break indicator

nSubsW <- length(waitDataLong)
nSubsC <- length(cogDataLong)
nSubsP <- length(physDataLong)
nSubsE <- length(easyDataLong)
totalSubs <- (nSubsW + nSubsC + nSubsP + nSubsE)

#  short lists
cogData <- list()
waitData <- list()
physData <- list()
easyData <- list()

# For pre-post analysis
cogBefore <- list()
cogAfter <- list()
waitBefore <- list()
waitAfter <- list()
physBefore <- list()
physAfter <- list()
easyBefore <- list()
easyAfter <- list()

# clean wait group data
for (subject in seq(nSubsW)){

  # Before/after
  breakW <- which(waitDataLong[[subject]]$V1 == 0)

  waitBefore[[subject]] <- waitDataLong[[subject]][1:(breakW - 1),]
  waitAfter[[subject]] <- waitDataLong[[subject]][(breakW + 1):length(waitDataLong[[subject]]$V1),]

  # assign header names
  colnames(waitBefore[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")
  colnames(waitAfter[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

  # Overall data cleaning
  tempW <- waitDataLong[[subject]]$V1 != 0

  # create a new list with only decision trials / without break
  waitData[[subject]] <- waitDataLong[[subject]][tempW,]

  # assign header names
  colnames(waitData[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

# clean cognitive group data
for (subject in seq(nSubsC)){

  # Before/After
  breakC <- which(cogDataLong[[subject]]$V1 == 0)

  cogBefore[[subject]] <- cogDataLong[[subject]][1:(breakC-1),]
  temp <- cogBefore[[subject]]$V7 # trial outcome vector
  quitTrials <- grep("Quit",temp)
  completeTrials <- grep("Reward",temp)
  decisionTrials <- sort(c(quitTrials,completeTrials))
  cogBefore[[subject]] <- cogBefore[[subject]][decisionTrials,]

  cogAfter[[subject]] <- cogDataLong[[subject]][(breakC+1):length(cogDataLong[[subject]]$V1),]
  temp <- cogAfter[[subject]]$V7 # trial outcome vector
  quitTrials <- grep("Quit",temp)
  completeTrials <- grep("Reward",temp)
  decisionTrials <- sort(c(quitTrials,completeTrials))
  cogAfter[[subject]] <- cogAfter[[subject]][decisionTrials,]

  colnames(cogBefore[[subject]]) <- c("Handling","Reward","Choice","Task RT","Choice RT",
                                      "Experiment Time","Trial Outcome","Task Type","Task Setup")
  colnames(cogAfter[[subject]]) <- c("Handling","Reward","Choice","Task RT","Choice RT",
                                     "Experiment Time","Trial Outcome","Task Type","Task Setup")

  # Overall data cleaning
  # create an index for the trials in which a decision was made (completed, quit) 
  temp <- cogDataLong[[subject]]$V7 # trial outcome vector
  quitTrials <- grep("Quit",temp)
  completeTrials <- grep("Reward",temp)
  decisionTrials <- sort(c(quitTrials,completeTrials))

  # create a new list with only decision trials / without break
  tempC <- cogDataLong[[subject]]$V1 != 0  
  cogData[[subject]] <- cogDataLong[[subject]][decisionTrials,]
  cogDataLong[[subject]] <- cogDataLong[[subject]][tempC,]

  colnames(cogData[[subject]]) <- c("Handling","Reward","Choice","Task RT","Choice RT",
                                    "Experiment Time","Trial Outcome","Task Type","Task Setup")
  colnames(cogDataLong[[subject]]) <- c("Handling","Reward","Choice","Task RT","Choice RT",
                                        "Experiment Time","Trial Outcome","Task Type","Task Setup") # Long contains within-trial info

# clean physical group data
for (subject in seq(nSubsP)){

  # Before/After
  breakP <- which(physDataLong[[subject]]$V1 == 0)

  physBefore[[subject]] <- physDataLong[[subject]][1:(breakP - 1),]
  physAfter[[subject]] <- physDataLong[[subject]][(breakP + 1):length(physDataLong[[subject]]$V1),]

  colnames(physBefore[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")
  colnames(physAfter[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

  # create a new list with only decision trials / without break  
  tempP <- physDataLong[[subject]]$V1 != 0
  physData[[subject]] <- physDataLong[[subject]][tempP,]
  colnames(physData[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

# clean pheaysical group data
for (subject in seq(nSubsE)){

  # Before/After
  breakE <- which(easyDataLong[[subject]]$V1 == 0)

  easyBefore[[subject]] <- easyDataLong[[subject]][1:(breakE - 1),]
  easyAfter[[subject]] <- easyDataLong[[subject]][(breakE + 1):length(easyDataLong[[subject]]$V1),]

  colnames(easyBefore[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")
  colnames(easyAfter[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

  # create a new list with only decision trials / without break  
  tempE <- easyDataLong[[subject]]$V1 != 0
  easyData[[subject]] <- easyDataLong[[subject]][tempE,]
  colnames(easyData[[subject]]) <- c("Handling","Reward","Choice","RT","Experiment Time")

# Combine before/after across subjects
cBefore <- rbind(cogBefore[[1]],cogBefore[[2]]) 
cAfter <- rbind(cogAfter[[1]],cogAfter[[2]]) 
wBefore <- rbind(waitBefore[[1]],waitBefore[[2]]) 
wAfter <- rbind(waitAfter[[1]],waitAfter[[2]]) 
pBefore <- rbind(easyBefore[[1]],physBefore[[2]])
pAfter <- rbind(physAfter[[1]],physAfter[[2]])
eBefore <- rbind(easyBefore[[1]],easyBefore[[2]])
eAfter <- rbind(easyAfter[[1]],easyAfter[[2]])

for (i in 3:nSubsW){

    wBefore <- rbind(wBefore,waitBefore[[i]])
    wAfter <- rbind(wAfter,waitAfter[[i]])

for (i in 3:nSubsC){

    cBefore <- rbind(cBefore,cogBefore[[i]])  
    cAfter <- rbind(cAfter,cogAfter[[i]])

for (i in 3:nSubsP){

    pBefore <- rbind(pBefore,physBefore[[i]])
    pAfter <- rbind(pAfter,physAfter[[i]])

for (i in 3:nSubsE){

    eBefore <- rbind(eBefore,easyBefore[[i]])
    eAfter <- rbind(eAfter,easyAfter[[i]])

#rm(cogBefore,cogAfter,waitBefore,waitAfter)

# aggregate ADD PHYS
waitBeforeAll <- aggregate(wBefore$Choice,
          by = list(wBefore$Reward, wBefore$Handling),
          FUN = 'mean')
waitAfterAll <- aggregate(wAfter$Choice,
          by = list(wAfter$Reward, wAfter$Handling),
          FUN = 'mean')
cogBeforeAll <- aggregate(cBefore$Choice,
          by = list(cBefore$Reward, cBefore$Handling),
          FUN = 'mean')
cogAfterAll <- aggregate(cAfter$Choice,
          by = list(cAfter$Reward, cAfter$Handling),
          FUN = 'mean')
physBeforeAll <- aggregate(pBefore$Choice,
          by = list(pBefore$Reward, pBefore$Handling),
          FUN = 'mean')
physAfterAll <- aggregate(pAfter$Choice,
          by = list(pAfter$Reward, pAfter$Handling),
          FUN = 'mean')
easyBeforeAll <- aggregate(eBefore$Choice,
          by = list(eBefore$Reward, eBefore$Handling),
          FUN = 'mean')
easyAfterAll <- aggregate(eAfter$Choice,
          by = list(eAfter$Reward, eAfter$Handling),
          FUN = 'mean')

rm(cBefore,cAfter,wBefore,wAfter, pBefore, pAfter, eBefore, eAfter)

# remove unnecessary variables
rm(temp,tempW,tempC,quitTrials,completeTrials,decisionTrials,subject)
```

``` {r Basic Prepping, echo=FALSE}  
# params for reference
handling <- c(2, 10, 14)
cond <- c("Wait", "Cognitive", "Physical", "Easy")

# create data frame with summary stats
tempW <- summarizeData()
tempC <- summarizeData(Data = cogData, nSubs = nSubsC, type = "Cognitive")
tempP <- summarizeData(Data = physData, nSubs = nSubsP, type = "Physical")
tempE <- summarizeData(Data = easyData, nSubs = nSubsE, type = "Easy")

summaryData <- rbind(tempW,
                     tempC,
                     tempP,
                     tempE)

colnames(summaryData) <- c("Group", "propComplete", "Earnings", "RT", "forcedTravel")

summaryData$Cols <- c(rep(cols[1],nSubsW), rep(cols[2],nSubsC), rep(cols[3],nSubsP), rep(cols[4],nSubsE))
rm(tempW, tempC, tempP, tempE)

# Indexes for groups
indxWait <- seq(nSubsW)
indxWaitLong <- seq(nSubsW * 9) # 9 being the number of combinations
indxCog <- (nSubsW + 1):(nSubsW + nSubsC)
indxCogLong <- ((nSubsW * 9) + 1):((nSubsW * 9) + (nSubsC * 9))
indxPhys <- ((nSubsW + nSubsC) + 1):(nSubsW + nSubsC + nSubsP)
indxPhysLong <- ((nSubsW * 9) + (nSubsC * 9) + 1):((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9))
indxEasy <- ((nSubsW + nSubsC + nSubsP) + 1):totalSubs #(nSubsW + nSubsC + nSubsP + nSubsE)
indxEasyLong <- ((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9) + 1):((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9) + (nSubsE * 9))

# Get the proportion completed per combination for GLMs
tempW <- propComplete()
tempC <- propComplete(Data = cogData, nSubs = nSubsC, indx = indxCog, type = "Cognitive")
tempP <- propComplete(Data = physData, nSubs = nSubsP, indx = indxPhys, type = "Physical")
tempE <- propComplete(Data = easyData, nSubs = nSubsE, indx = indxEasy, type = "Easy")

allProp2 <- rbind(tempW,
                  tempC,
                  tempP,
                  tempE)  
allProp2$totalComplete <- allProp2$propComplete * allProp2$totalTrials

rm(tempW, tempC, tempP, tempE)

# add optimal vector and difference from optimal
wait <- c(rep(1,nSubsW*3),
          c(rep(0,nSubsW), rep(1,nSubsW*2)),
          c(rep(0,nSubsW*2), rep(1,nSubsW)))
cog <- c(rep(1,nSubsC*3),
          c(rep(0,nSubsC), rep(1,nSubsC*2)),
          c(rep(0,nSubsC*2), rep(1,nSubsC)))
phys <- c(rep(1,nSubsP*3),
          c(rep(0,nSubsP), rep(1,nSubsP*2)),
          c(rep(0,nSubsP*2), rep(1,nSubsP)))
easy <- c(rep(1,nSubsE*3),
          c(rep(0,nSubsE), rep(1,nSubsE*2)),
          c(rep(0,nSubsE*2), rep(1,nSubsE)))
allProp2$Optimal <- c(wait,cog,phys,easy)
allProp2$Deviation <- allProp2$Optimal - allProp2$propComplete

# sort by subject ID
#allProp2 <- allProp2[order(allProp2$SubjID),]

# actual proportions and standard errors, and the optimal summary for plotting
trueProp <- aggregate(allProp2$propComplete,
            by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
            FUN = 'mean')

trueSD <- aggregate(allProp2$propComplete,
          by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
          FUN = 'sd')
colnames(trueProp) <- c("Reward","Handling","Group","propComplete")
trueProp$SE[1:9] <- (trueSD$x[1:9])/(sqrt(nSubsW))
trueProp$SE[10:18] <- (trueSD$x[10:18])/(sqrt(nSubsC))
trueProp$SE[19:27] <- (trueSD$x[19:27])/(sqrt(nSubsP))
trueProp$SE[28:36] <- (trueSD$x[28:36])/(sqrt(nSubsE))

trueProp$Optimal <- aggregate(allProp2$Optimal,
            by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
            FUN = 'mean')$x
```

## Analyses

@@

``` {r 16.1.1., fig.align="center", fig.width=4, fig.height=4, echo=FALSE}
# Add an AR predictor based on the number of quits
waitData <- lapply(waitData, function(data) data %>% mutate(AR = seqQuits(Choice)))
cogData <- lapply(cogData, function(data) data %>% mutate(AR = seqQuits(Choice)))
physData <- lapply(physData, function(data) data %>% mutate(AR = seqQuits(Choice)))
easyData <- lapply(easyData, function(data) data %>% mutate(AR = seqQuits(Choice)))

# Logistic fits
waitLogis <- lapply(waitData, function(data) glm(Choice ~ Handling + Reward, data = data, family = "binomial"))
cogLogis <- lapply(cogData, function(data) glm(Choice ~ Handling + Reward, data = data, family = "binomial"))
physLogis <- lapply(physData, function(data) glm(Choice ~ Handling + Reward, data = data, family = "binomial"))
easyLogis <- lapply(easyData, function(data) glm(Choice ~ Handling + Reward, data = data, family = "binomial"))

# Coefficients
waitLogis_coeffs <- t(sapply(waitLogis, "[[", "coefficients"))
cogLogis_coeffs <- t(sapply(cogLogis, "[[", "coefficients"))
physLogis_coeffs <- t(sapply(physLogis, "[[", "coefficients"))
easyLogis_coeffs <- t(sapply(easyLogis, "[[", "coefficients"))
allLogis_coeffs <- rbind(waitLogis_coeffs,
                         cogLogis_coeffs,
                         physLogis_coeffs,
                         easyLogis_coeffs)

# Rank sums
rankHand <- wilcox.test(allLogis_coeffs[,2])
rankReward <- wilcox.test(allLogis_coeffs[,3])

# Plot coeffs
temp <- melt(allLogis_coeffs[,2:3])
qplot(data = temp, x = Var2, y = value, geom = "boxplot") + 
  labs(title = "Handling and Reward Coefficients", x = "", y = "Coefficients") +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme_classic()

rm(waitLogis, cogLogis, physLogis, easyLogis)

@@

``` {r 16.1.2., echo=FALSE, fig.align="center"}
# Logistic fits
waitLogis <- lapply(waitData, function(data) glm(Choice ~ Handling + Reward + AR, data = data, family = "binomial"))
cogLogis <- lapply(cogData, function(data) glm(Choice ~ Handling + Reward + AR, data = data, family = "binomial"))
physLogis <- lapply(physData, function(data) glm(Choice ~ Handling + Reward + AR, data = data, family = "binomial"))
easyLogis <- lapply(easyData, function(data) glm(Choice ~ Handling + Reward + AR, data = data, family = "binomial"))

# Coefficients
waitLogis_coeffs <- t(sapply(waitLogis, "[[", "coefficients"))
cogLogis_coeffs <- t(sapply(cogLogis, "[[", "coefficients"))
physLogis_coeffs <- t(sapply(physLogis, "[[", "coefficients"))
easyLogis_coeffs <- t(sapply(easyLogis, "[[", "coefficients"))
allLogis_coeffs <- rbind(waitLogis_coeffs,
                         cogLogis_coeffs,
                         physLogis_coeffs,
                         easyLogis_coeffs)

#waitLogis_coeffs <- sapply(waitLogis, function(data) (confint(data)[4,1] < 0 & confint(data)[4,2] > 0))

# Rank sums
rankAR <- wilcox.test(allLogis_coeffs[,4])

# Plot coeffs
temp <- melt(allLogis_coeffs[,2:4])

# The ugliest code, to check the existence of 0 on the CIs
AR_CIs <- data.frame(Subj = NULL, Low = NULL, High = NULL)
temp <- data.frame(Subj = seq(nSubsW), Low = rep(-0.1,nSubsW), High = rep(0,nSubsW))
for (i in seq(length(waitLogis))) {
  test <- tryCatch(round(confint(waitLogis[[i]]), digits = 2)[4,], error=function(e){})
  if(!is.null(test)) {
    temp$Low[i] <- test[1]
    temp$High[i] <- test[2]
  }
}
AR_CIs <- rbind(AR_CIs, temp)
temp <- data.frame(Subj = seq(nSubsC), Low = rep(0,nSubsC), High = rep(0,nSubsC))
for (i in seq(length(cogLogis))) {
  test <- tryCatch(round(confint(cogLogis[[i]]), digits = 2)[4,], error=function(e){})
  if(!is.null(test)) {
    temp$Low[i] <- test[1]
    temp$High[i] <- test[2]
  }
}
AR_CIs <- rbind(AR_CIs, temp)
temp <- data.frame(Subj = seq(nSubsP), Low = rep(0,nSubsP), High = rep(0,nSubsP))
for (i in seq(length(physLogis))) {
  test <- tryCatch(round(confint(physLogis[[i]]), digits = 2)[4,], error=function(e){})
  if(!is.null(test)) {
    temp$Low[i] <- test[1]
    temp$High[i] <- test[2]
  }
}
AR_CIs <- rbind(AR_CIs, temp)
temp <- data.frame(Subj = seq(nSubsE), Low = rep(0,nSubsE), High = rep(0,nSubsE))
for (i in seq(length(easyLogis))) {
  test <- tryCatch(round(confint(easyLogis[[i]]), digits = 2)[4,], error=function(e){})
  if(!is.null(test)) {
    temp$Low[i] <- test[1]
    temp$High[i] <- test[2]
  }
}
AR_CIs <- rbind(AR_CIs, temp)
AR_CIs$High[is.na(AR_CIs$High)] <- 0.1
AR_CIs$AR_Effect <- apply(AR_CIs,1, function(x) !(x[2] <= 0 & x[3] >= 0))

# Around 10% seemed to have been influenced by previous quitting
propAR_effects <- (sum(AR_CIs$AR_Effect) / totalSubs) * 100

rm(temp,waitLogis, cogLogis, physLogis, easyLogis)

To measure the significance of each subject's autoregressive coefficient, we computed its CI and checked how many contained zero. By this measure, around `r round(propAR_effects)` percent of our participants seemed to have been influenced by recent quitting. However, the autoregressive predictor did not preclude the effect of the remaininig experimental parameters.

@@

``` {r 16.2.1. Overall Proportion/Earnings,fig.align="center",fig.width=10,fig.height=4, echo=FALSE}  
## prop complete
# ecdfs
wait <- ecdf(summaryData$propComplete[indxWait])
cog <- ecdf(summaryData$propComplete[indxCog])
phys <- ecdf(summaryData$propComplete[indxPhys])
easy <- ecdf(summaryData$propComplete[indxEasy])
range <- seq(0,1,by=0.01)

test <- data.frame(range = range,
                   ecdf = c(wait(range),cog(range),phys(range),easy(range)),
                   Group = factor(c(rep("Wait",101),rep("Cognitive",101),rep("Physical",101),rep("Easy", 101)), levels = c("Wait","Cognitive","Physical","Easy")))

p1 <- ggplot(test, aes(x=range,y=ecdf, group=Group, color=Group))
p1 <- p1 + geom_line(show.legend = F, size = 1.5) +
    scale_color_manual(values = cols) +
    labs(title = "ECDF of Acceptances per Group", y="ECDF", x = "Proportion Accepted") +
    theme_classic()

# problem is, data is not normal, so we're going to bootstrap it and DEFINITELY permute
# Bootstrap nonparametric
B <- 5000
meanBootW <- bootstrap(summaryData$propComplete[indxWait],B = B) #rep(0,B)
meanBootC <- bootstrap(summaryData$propComplete[indxCog], B = B) #rep(0,B)
meanBootP <- bootstrap(summaryData$propComplete[indxPhys], B = B)
meanBootE <- bootstrap(summaryData$propComplete[indxEasy], B = B)

#CIs wait
meanW <- mean(meanBootW)
upperW <- quantile(meanBootW,0.975) # meanBoot?
lowerW <- quantile(meanBootW,0.025)

#CIs effort
meanC <- mean(meanBootC)
upperC <- quantile(meanBootC,0.975) # meanBoot?
lowerC <- quantile(meanBootC,0.025)

#CIs physical
meanP <- mean(meanBootP)
upperP <- quantile(meanBootP,0.975) # meanBoot?
lowerP <- quantile(meanBootP,0.025)

#CIs easy
meanE <- mean(meanBootE)
upperE <- quantile(meanBootE,0.975) # meanBoot?
lowerE <- quantile(meanBootE,0.025)

# formal testing 
# proportion completed
propAov <- aov(propComplete ~ factor(Group), data = summaryData)

# effect size (get the one for anovas)
# propRsq <- cohenD(summaryData$propComplete[indxWait],summaryData$propComplete[indxCog])

# post-hoc
temp <- do.call(cbind, lapply(split(summaryData, summaryData$Group),"[[","propComplete"))
temp2 <- t(combn(colnames(temp),2))

propTests <- data.frame(Comparison = paste(temp2[,1], temp2[,2]),
                        T_pval = rep(1,6),
                        CohenD = rep(0,6))

for (i in seq(nrow(temp2))) {

  propTests$T_pval[i] <- round(permute(temp[,temp2[i,1]], temp[,temp2[i,2]])$Pval, digits = 2)
  propTests$CohenD[i] <- round(cohenD(temp[,temp2[i,1]], temp[,temp2[i,2]]), digits = 2)

# earnings
earnAov <- aov(Earnings ~ factor(Group), data = summaryData)

# effect size
# earnRsq <- cohenD(summaryData$propComplete[indxWait],summaryData$propComplete[indxCog])

# post-hoc
temp <- do.call(cbind, lapply(split(summaryData, summaryData$Group),"[[","Earnings"))
temp2 <- t(combn(colnames(temp),2))

earnTests <- data.frame(Comparison = paste(temp2[,1], temp2[,2]),
                        T_pval = rep(1,6),
                        CohenD = rep(0,6))

for (i in seq(nrow(temp2))) {

  earnTests$T_pval[i] <- round(permute(temp[,temp2[i,1]], temp[,temp2[i,2]])$Pval, digits = 2)
  earnTests$CohenD[i] <- round(cohenD(temp[,temp2[i,1]], temp[,temp2[i,2]]), digits = 2)

>>>>

### 4.2.

Prospect acceptance rates will differ across four between-subject conditions, in which the delays associated with rewards (a) are unfilled, (b) include a cognitive effort requirement, (c) include a physical effort requirement, or (d) require a trivial level of physical effort but have matched visual stimuli to the physical effort condition.

  **4.2.1.**	In the cognitive effort condition, overall acceptance rates will be greater than in the unfilled-delay condition.

@@

### 4.3.

Choices will be well fit by a computational model in which the subjective opportunity cost of time is free to vary across the four between-subject conditions.

  **4.3.1.**	Participants will display stable preferences, meaning that the reward amounts they accept in a given timing condition will be similar throughout the experiment.

@@

## 5. Existing data

Registration prior to creation of data

## 6. Data collection

The study will use human subjects recruited from the university community, who will be recruited through the university job board. We will pay up to $30 for a session (1 hour), which includes incentive-based rewards. Participants will be ineligible if they have participated in our pilot study, or they are under 18 years of age. The study will run for however long it takes to reach our desired sample size.

## 7. Sample size

We will acquire a sample of 48 participants.

## 8. Sample size rationale?

Sample sizes were determined by means of power analysis (repeated measures ANOVA), using a significance level of 0.05, power of 0.8, a desired effect size of f = 0.5, and four groups (one for each condition). The resulting sample size was 45, which we increased to 48 in order to match the potential order of blocks and colors (section 15) while further ensuring that we have enough power.

## 9. Variables

### 9.1 The present study has three levels of manipulation.

9.1.1. Cost condition, manipulated within subjects. The conditions are (a) unfilled delay, (b) cognitive effort, and (c) physical effort. Unfilled delays will be paired with each effort type in a blocked manner (7-minute blocks). Each combination will be experienced three times.

9.1.2. Handling and travel times. These will dictate how long each trial is. The "handling time" governs how long it is necessary to wait or work to obtain a reward if a trial is accepted. The "travel time" governs how long the inter-trial interval will be, regardless of whether the trial is accepted. Participants will know that the handling time will be always 10 seconds, and the travel time 6 seconds (matching one of the between-group experiment conditions).

9.1.3. Reward magnitude, manipulated within-subject across individual trials. On each trial, participants will learn the amount of money they can earn if they complete the trial (i.e. 4, 8, or 20 cents). The amounts will be sampled uniformly, while guaranteeing that each participant experiences each amount at least twice during each block.

## 10. Measured variables

### 10.1. Behavioral variables

10.1.1. Choice (complete or quit a given trial).

10.1.2. Response times.

10.1.3. Total earnings.

10.1.3. Performance in the cognitive effort tasks (right or wrong answer).

10.1.4. Grip force throughout the experiment and maximum grip strength (physical effort condition).

### 10.2. Questionnaire

10.2.1. Age.

10.2.2. Gender.

10.2.3. Post-experimental oral assessment of choice strategy adoption. We will ask participants the following: "Did you follow a specific acceptance strategy? If so, what was it?" and "Which condition felt most effortful to you?". Answers will be written down by the experimenter, and recorded on an excel spreadsheet.

## 11. Indices

### 11.1. Behavioral data

11.1.1. Proportion of prospects accepted.

11.1.2. Percentage of cognitive effort trials with above-threshold errors (over 2 errors), in which participants are forced to travel to the next trial.

### 11.2. Optimal behavior

The theoretical opportunity cost of time will be calculated for each handling/travel/reward acceptance combination in order to assess choice optimality. This measure is computed as the cents per second possible when consistently accepting only certain reward amounts (e.g. 8 and 20 cents) in a given timing condition. This ground truth measure is independent of cost type, allowing us to assess subject optimality.

## 12. Study type

Experiment - A researcher randomly assigns treatments to study subjects, this includes field or lab experiments. This is also known as an intervention experiment and includes randomized controlled trials.

## 13. Blinding

No blinding is involved in this study.

## 14. Study design

The present study complements our previous one, which evaluated cost types through a between-group design. Many of the details remain the same. However, this time every participant will experience all forms of cost (delay, physical effort, and cognitive effort). Delay trials will be intermixed with both physical and cognitive trials, providing a common condition to compare these effortful costs. On each trial, participants will have the opportunity to either wait or work for 10 seconds (the "handling time") in order to earn a reward; alternatively, they can quit by pressing the space bar and skip to the next trial (which incurs a 6 second travel time). Two types of trial combinations (physical/wait and cognitive/wait) will be experienced three times by each participant in 7-minute-long, interleaved blocks. Reward amounts can be 4, 8, or 20 cents, uniformly distributed. Timing information will be disclosed at the beginning of each block, and reward prospects displayed before each trial begins.

## 15. Randomization

### 15.1. Block type 
Each block will contain unfilled delay trials and trials of one effortful cost type (se 14.1). Blocks with each cost type will be interleaved, and participants will be assigned to one of two orders: beginning with a cognitive/wait or a physical/wait block. A row will be assigned to subject IDs prior to recruitment, ensuring that each order is experienced equally often. Participants will be offered a break after completing three out of the six blocks.

### 15.2. Rewards and cost type 
Participants can earn 4, 8, or 20 cents per trial at the cost of performing one of the three conditions. These values are sampled without replacement from a vector containing every combination of cost (effort defined by the current block, and delay) and reward. The vector is reshuffled once completed. This ensures that each combination of cost and reward is presented at least twice per block.

### 15.3. Trial color 
During the offer phase, each cost type will appear in a unique color in order to ease its identification. Colors were chosen to contrast each other while being amenable to red-green color-blindness. Colors will be assigned to each cost type so that all three colors are equally paired to each cost type across participants. A row will be assigned to subject IDs prior to recruitment to indicate their block order and color pairing.

## 16. Analyses

@@

``` {r Load Data, echo=FALSE}
### data loading and cleaning up
setwd("./data")
files <- dir(pattern = '_log.csv')
nSubjs <- length(files)

dataAll <- tibble(SubjID = files) %>%
  mutate(contents = map(SubjID, ~ read_csv(., col_types = cols())))  %>%
  mutate(Cost = substring(SubjID, 5, 8),
         Cost = case_when(Cost == "wait" ~ "Wait",
                          Cost == "cogT" ~ "Cognitive",
                          Cost == "phys" ~ "Physical",
                          Cost == "phea" ~ "Easy"),
         SubjID = as.integer(substring(SubjID, 0, 3))) %>%
  unnest() %>%
  mutate(Half = ifelse(Block < 4, "Half_1", "Half_2"),
         Cost = factor(Cost, levels = c("Physical", "Cognitive", "Wait", "Easy")),
         optimal = case_when(
          (Handling == 10 & Reward < 8) ~ 0,
          (Handling == 14 & Reward < 20) ~ 0,
          TRUE ~ 1
         ))

# get a simple subject list and the number of subjects
subjList <- unique(dataAll$SubjID)
nSubj <- length(subjList)

@@

``` {r 16.1.1., fig.align="center", fig.width = 4, fig.height = 4, echo = FALSE}
# filterout errors and compute R + H logistic per subject, then get coefficients
logisRH <- dataAll %>%
  filter(Choice < 2) %>%
  plyr::dlply("SubjID", identity) %>%
  lapply(function(data) {glm(Choice ~ Handling + Reward, data = data, family = "binomial")}) %>%
  sapply(coefficients)

# Rank sum tests
rankHand <- wilcox.test(logisRH["Handling", ])
rankReward <- wilcox.test(logisRH["Reward", ])

# Plot coeffs
temp <- melt(logisRH[c("Handling", "Reward"), ])
qplot(data = temp, x = Var1, y = value, geom = "boxplot") + 
  labs(y = "Coefficients") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(legend.position = c(0.9, 0.7),
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.background = element_blank(), 
       axis.line = element_line(colour = "black"),
       text = element_text(size = 16))
```

@@

``` {r 16.1.2., echo = FALSE, include = FALSE}
# add an AR regressor to track recent quits
logisRHAR <- dataAll %>%
  filter(Choice < 2) %>%
  group_by(SubjID) %>%
  mutate(AR = seqQuits(Choice)) %>%
  plyr::dlply("SubjID", identity) %>%
  lapply(function(data) {glm(Choice ~ Handling + Reward + AR, data = data, family = "binomial")})

# Rank sums
rankAR <- logisRHAR %>% 
  sapply(coefficients) %>%
  wilcox.test(.["AR", ])

AR_CIs <- sapply(logisRHAR, function(x) {tryCatch(confint(x)[4, ], error = function(e){c(0, 0)})}) %>%
  replace_na(0) %>%
  t() %>% 
  apply(1, function(x) !(x[1] <= 0 & x[2] >= 0))

# Around 10% seemed to have been influenced by previous quitting
prcntAR_effects <- mean(AR_CIs) * 100
```

To measure the significance of each subject's autoregressive coefficient, we computed its CI and checked how many contained zero. By this measure, around `r round(prcntAR_effects)` percent of our participants seemed to have been influenced by recent quitting. However, the autoregressive predictor did not preclude the effect of the remaininig experimental parameters.

@@

``` {r 16.2.1. Overall Proportion/Earnings, fig.align = "center", fig.width = 8, fig.height = 4, echo = FALSE}  
## ECDF of proportion completed per cost type
temp <- dataAll %>%
        filter(Choice < 2) %>%
        mutate(Cost = factor(Cost, levels = list("Physical", "Cognitive", "Wait", "Easy"))) %>%
        group_by(SubjID, Cost) %>%
        summarise(Earnings = sum(Reward[Choice == 1] / 100),
                  pComplete = mean(Choice))

## formal testing 
# proportion completed
propAov <- aov(pComplete ~ factor(Cost), data = temp)

propTests <- temp %>% 
  plyr::dlply("Cost", identity) %>%
  sapply("[[", "pComplete") %>%
  as.data.frame() %>%
  combn(2, simplify = F) %>%
  sapply(function(x) {
    c(contrast = paste(colnames(x), collapse = "vs"), 
      permPval = round(permute(x[, 1], x[, 2], simple = T), digits = 3),
      cohenD = round(cohenD(x[, 1], x[, 2]), digits = 3))
  }) %>% 
  t()

# pairwise.wilcox.test(temp$pComplete, temp$Cost)

# earnings
earnAov <- aov(Earnings ~ factor(Cost), data = temp)

earnTests <- temp %>% 
  plyr::dlply("Cost", identity) %>%
  sapply("[[", "Earnings") %>%
  as.data.frame() %>%
  combn(2, simplify = F) %>%
  sapply(function(x) {
    c(contrast = paste(colnames(x), collapse = "vs"), 
      permPval = round(permute(x[, 1], x[, 2], simple = T), digits = 3),
      cohenD = round(cohenD(x[, 1], x[, 2]), digits = 3))
  }) %>% 
  t()

# plots
p1 <- ggplot(aes(pComplete, color = Cost), data = temp) +
        stat_ecdf(lwd = 1.2, show.legend = F) +
        scale_color_manual(values = cols) +
        xlim(0, 1) +
        labs(y="ECDF", x = "Proportion Accepted") +
        theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              panel.background = element_blank(), 
              axis.line = element_line(colour = "black"),
              text = element_text(size = 16))

p2 <- ggplot(temp, aes(Cost, pComplete, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        ylim(c(0, 1)) +
        labs(y = "Proportion Completed") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

p3 <- ggplot(temp, aes(Cost, Earnings, fill = Cost)) + 
        geom_boxplot(show.legend = F) +
        scale_fill_manual(values = cols) +
        labs(y = "Total Earnings (dollars)") +
        theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(), 
            panel.background = element_blank(), 
            axis.line = element_line(colour = "black"),
            text = element_text(size = 16))

grid.arrange(p2, p3, ncol = 2)
```
